---
title: "Calculate block information"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{calc_blockinfo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette shows how to calculate block information and block information summaries for an empirical sample.

```{r setup}
library(MFCblockInfo)
```

Load the data.
```{r}
data("data.mt")
str(data.mt)
```
## Small working example
Computationally feasible: only for the first two persons, and blocks 1 and 3.
```{r}
info1 <- calc.info.block(lhb.mplus, traits=as.matrix(data.mt$traits[1:2,-ncol(data.mt$traits)]),
                         int=data.mt$int, loads=data.mt$loads, uni=data.mt$uni, K=data.mt$K, nb=data.mt$nb,
                         which.blocks=c(1,3))
```

Obtain standard errors / estimation variances, sum information across all blocks (summed=TRUE), return variances (var.out=TRUE)
```{r}
ses1 <- info2se(info1, summed=T, var.out=T)
```

Calculate the trace of the information matrix, average across persons
```{r}
trace1 <- calc.info.trace(info1, avg=F)
```

## Example with full dataset
(To save computational time, the stored result can be used.)
```{r}
# info.mt <- calc.info.block(lhb.mplus, traits=as.matrix(data.mt$traits[,-ncol(data.mt$traits)]),
#                            int=data.mt$int, loads=data.mt$loads, uni=data.mt$uni,
#                            K=data.mt$K, nb=data.mt$nb) #which.blocks defaults to all
info.mt <- readRDS("data/info_all_emp_MT.rds")
```

Calculate and examine block $R^2$ for all blocks, averaged across persons.
```{r}
r2.mt.all <- do.call(rbind, lapply(1:20, function(b, info.emp) colMeans(calc.info.block.r2(info.emp, wo.blocks=b))
                                   , info.emp=info.mt))
round(r2.mt.all, 2)
psych::describe(r2.mt.all)
```

Calculate the trace of test information (T-optimality), averaged across persons.
```{r}
trace.mt.all <- calc.info.trace(info.mt, avg=T)
```

Calculate and examine the test information for the empirical sample.
```{r}
ses.mt <- info2se(info.mt, summed=T)

#scatterplot
par(mfrow=c(1,2))
for(tr in 1:ncol(ses.mt)) {
  plot(data.mt$traits[,tr], ses.mt[,tr], main=paste("Trait",tr),
       xlab="MT", ylab="SE")
}
round(colMeans(ses.mt), 2)
```

## Block information plots

Block information, plot blocks 3 and 4.
```{r}
# info.mt.plots <- calc.info.plot(tr.levels=seq(-2,2,.5), fix.levels=c(-1,0,1), fix.level.others=0, K=data.mt$K, which.blocks=1:data.mt$K,
#                           lhb.mplus,
#                           int=data.mt$int, loads=data.mt$loads, uni=data.mt$uni, nb=data.mt$nb)
# saveRDS(info.mt.plots, file="data/info_mt_plots.rds")
info.mt.plots <- readRDS("data/info_mt_plots.rds")
plot.block(which.blocks = 3:4, info=info.mt.plots, K=20, loads=data.mt$loads)
```

Test information (\textit{SE}s).
```{r}
# testinfo.mt.plots <- calc.testinfo.plot(tr.levels=seq(-2,2,.5), fix.level.others=0, K=data.mt$K, which.blocks=1:data.mt$K,
#                                         FUN=lhb.mplus,
#                                         int=data.mt$int, loads=data.mt$loads, uni=data.mt$uni, nb=data.mt$nb)
# saveRDS(testinfo.mt.plots, file="data/testinfo_mt_plots.rds")
testinfo.mt.plots <- readRDS("data/testinfo_mt_plots.rds")
plot.testinfo(testinfo.mt.plots, loads=data.mt$loads, par.mfrow=c(2,5))
```

only 1 trait varied, others randomly drawn from MVN, n for each trait level
```{r}
# testinfo.mt.1d <- calc.testinfo.1d(tr.levels=seq(-2,2,.5), sigma=cor(data.mt$traits[,1:5]), n=10, which.blocks = 1:data.mt$K,
#                                    seed=42,
#                                    FUN=lhb.mplus,
#                                    int=data.mt$int, loads=data.mt$loads, uni=data.mt$uni,
#                                    K=data.mt$K, nb=data.mt$nb)
#plot with base r or anything else
# for(tr in 1:ncol(testinfo.mt.1d$ses)) plot(testinfo.mt.1d$variedlevels, testinfo.mt.1d$ses[,tr], type="l")
```

